---
title: "CC-RIX Transcript Imputation Check"
author: Anna L Tyler
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    collapsed: no
    toc: yes
    toc_float: yes
---

## Introduction

```{r get_args}
args <- commandArgs(trailingOnly = T)
tissue.name <- args[1]

if(is.na(tissue.name)){
    tissue.name <- "Adipose"
}

is.interactive <- FALSE
#is.interactive <- TRUE
```

The purpose of this workflow is to compare `r tissue.name` transcripts 
imputed in the CC to the measured `r tissue.name` transcripts.


```{r source_code}
library("here")

all.fun <- list.files(here("Code"), pattern = ".R", full.names = TRUE)
for(j in 1:length(all.fun)){source(all.fun[j])}
```


```{r load_libraries,  warning = FALSE, message = FALSE, error = FALSE}
needed.packages <- c("pheatmap", "knitr", "qtl2")
load_libraries(needed.packages, personal.library = TRUE)
```

Read in measured CC-RIX gene expression. Also read in DO 
data for comparison.

```{r read_data}
all.var <- ls()
data.loaded <- as.logical(length(which(all.var == "dataset.DO.Cube.Liver")))
cc.dir <- here("Data", "CC-RIX")

if(!data.loaded){
    annot <- read.csv(file.path(cc.dir, "final_sample_annotations.csv"))
    tissue.expr <- readRDS(file.path(cc.dir, "Expression.by.Tissue.RDS"))
    manifest <- as.matrix(read.csv(file.path(cc.dir, "Original_Samples_Manifest.csv")))
    covar <- as.matrix(read.csv(file.path(cc.dir, "Covar.csv"), row.names = 1))
    groups <- read.csv(file.path(cc.dir, "Groups.csv"), row.names = 1)
    group.def <- colnames(groups)
    load(here("Data", "dataset.DO.CUBE.multissue.RData")) #DO data for eQTL LOD scores
    data.set.name <- paste0("dataset.DO.Cube.", tissue.name)
    gene.table <- as.matrix(get(data.set.name)$annot.mrna)
}

do.expr <- get(paste0("dataset.DO.Cube.", tissue.name))
do.covar <- do.expr$covar.matrix
expr.annot <- do.expr$annot.mrna
expr.lod <- do.expr$lod.peaks
split.lod.markers <- strsplit(expr.lod$additive[,"marker.id"], "_")
lod.marker.chr <- sapply(split.lod.markers, function(x) x[1])
lod.marker.pos <- as.numeric(sapply(split.lod.markers, function(x) x[2]))
bp.buffer <- 4e6 #cis is anything within 4 Mb of the nearest marker

```

Adjust measured expression for all covariates: RNASeq_Batch, Age, 
Sex, Diet, and Treatment.

```{r adjust}
tissue.locale <- grep(tissue.name, names(tissue.expr), ignore.case = TRUE)

orsam.id <- names(rownames(tissue.expr[[tissue.locale]]))
orsam.idx <- match(orsam.id, rownames(covar))

tissue.info <- unique(t(sapply(rownames(tissue.expr[[tissue.locale]]), 
  function(x) get_mouse_info(x, manifest, "Barcode.1"))))

group.mat <- tissue.info[,group.def]
rnaseq.covar <- dummy_covar(covar[orsam.idx,c("Age", "RNAseq_Batch", colnames(group.mat))])
rownames(rnaseq.covar) <- tissue.info[,1]
cc.expr <- tissue.expr[[tissue.locale]]
adj.cc.expr <- adjust(cc.expr, rnaseq.covar)
do.adj.expr <- adjust(do.expr$data$norm, do.expr$covar.matrix)
```

```{r cc_data}
match_geno_pheno <- function(mouse.genoprobs, mouse.K, mouse.pheno.id, manifest){
    #get mouse info from the mouse id
    all.pheno.id <- t(sapply(mouse.pheno.id, function(x) get_CC_RIX_mouse_info(x, manifest)))

    all.geno.group <- apply(all.pheno.id, 1, function(x) paste(gsub(" X ", "_", x[6]), x[3], sep = "_"))
    #match up with the genotype data
    all.pheno.idx <- match(all.geno.group, rownames(mouse.genoprobs[[1]]))
    #take out na's
    not.na.locale <- which(!is.na(all.pheno.idx))
    all.pheno.idx <- all.pheno.idx[not.na.locale]
    mouse.pheno.id <- mouse.pheno.id[not.na.locale]
    
    new.genoprobs <- mouse.genoprobs
    new.K <- mouse.K
    #expand genoprobs to incorporate all these individuals
    #assign unique IDs to all individuals
    for(i in 1:length(mouse.genoprobs)){
        new.genoprobs[[i]] <- mouse.genoprobs[[i]][all.pheno.idx,,]
        rownames(new.genoprobs[[i]]) <- as.vector(mouse.pheno.id)
        new.K[[i]] <- mouse.K[[i]][all.pheno.idx,all.pheno.idx]
        rownames(new.K[[i]]) <- colnames(new.K[[i]]) <- as.vector(mouse.pheno.id)
    }

    result <- list("expanded.geno" = new.genoprobs, "expanded.K" = new.K)
    return(result)
}

cc.genoprobs <- readRDS(file.path(cc.dir, "Genoprobs.RDS")) #created by 3.5.CC-RIX_genotypes.Rmd
cc.K <- readRDS(file.path(cc.dir, "Kinship.RDS"))
cc.map <- readRDS(file.path(cc.dir, "Map.RDS"))
expanded.geno.K <- match_geno_pheno(cc.genoprobs, cc.K, rownames(cc.expr), manifest)
expanded.geno <- expanded.geno.K$expanded.geno
expanded.K <- expanded.geno.K$expanded.K
```


Read in imputed CC-RIX gene expression and variance components.
The variance components tell us how much variance was explained
by local genotype, distal genotype, and covariates for each
transcript in the DO tissues.

```{r load_imputed}
imp.cc.expr <- readRDS(here("Data", "imputed", 
    paste0("CC-RIX_", tissue.name, "_imputed_transcriptome_local.RDS")))

var.comp <- readRDS(here("Data", "imputed", 
    paste0("DO_", tissue.name, "_imputed_variance_decomposition.RDS")))
```

The distribution of variance explained by each component is
as follows:

```{r var_comp_dist}
boxplot(var.comp)
```

Group the measured expression based on the CC-RIX crosses.

```{r group_expr}
cross.id <- tissue.info[,"User.Defined.Strain"]
u_cross <- unique(cross.id)
cross.idx <- lapply(u_cross, function(x) which(cross.id == x))
cross_expr <- lapply(cross.idx, function(x) adj.cc.expr[x,])
names(cross_expr) <- u_cross
mean.cross.expr <- t(sapply(cross_expr, colMeans))
```

For each of these crosses pull out the imputed autosomal 
gene expression for comparison.

```{r imp_expr_by_strain}
split.cross <- strsplit(u_cross, " ")
mom <- sapply(split.cross, function(x) x[1])
dad <- sapply(split.cross, function(x) x[3])

#str(imp.expr)

mom.idx <- sapply(mom, function(x) grep(x, rownames(imp.cc.expr[[1]])))
dad.idx <- sapply(dad, function(x) grep(x, colnames(imp.cc.expr[[1]])))

#cbind(mom, rownames(imp.cc.expr[[1]])[mom.idx])
#cbind(dad, colnames(imp.cc.expr[[1]])[dad.idx])

imp.auto.expr <- t(sapply(1:length(mom.idx), function(x) imp.cc.expr[[1]][mom.idx[x], dad.idx[x],]))
rownames(imp.auto.expr) <- u_cross

#also make an expanded version to compare to adj.cc.expr
ind.cross <- tissue.info[,"User.Defined.Strain"]
split.cross <- strsplit(ind.cross, " ")
ind.mom <- sapply(split.cross, function(x) x[1])
ind.dad <- sapply(split.cross, function(x) x[3])

ind.mom.idx <- sapply(ind.mom, function(x) grep(x, rownames(imp.cc.expr[[1]])))
ind.dad.idx <- sapply(ind.dad, function(x) grep(x, colnames(imp.cc.expr[[1]])))

ind.auto.expr <- t(sapply(1:length(ind.mom.idx), function(x) imp.cc.expr[[1]][ind.mom.idx[x], ind.dad.idx[x],]))
rownames(ind.auto.expr) <- ind.cross
```

## Correlation between measured and imputed expression

Compare expression between the two data sets. The following 
histogram shows the distribution of correlations between
imputed gene expression and measured gene expression.

```{r match_transcripts}
common.transcripts <- intersect(colnames(adj.cc.expr), colnames(ind.auto.expr))
#test.transcript <- sample(common.transcripts, 1)
#plot.with.model(adj.cc.expr[,test.transcript], ind.auto.expr[,test.transcript],
#    xlab = "CC-RIX expression", ylab = "Imputed Expression")

all.cor <- sapply(common.transcripts, function(x) cor(adj.cc.expr[,x], ind.auto.expr[,x]))
med.cor <- median(all.cor, na.rm = TRUE)
hist(all.cor, breaks = 100);abline(v = med.cor, col = "red")
```

## Correlation between CC-RIX - DO correspondence and local eQTL LOD score

The following plot shows the correlation between the local additive 
eQTL LOD score for each transcript in the DOs and the correlation 
between the imputed and the actual gene expression.

```{r lod}

cis.lod <- rep(NA, length(common.transcripts))
names(cis.lod) <- common.transcripts
for(i in 1:length(common.transcripts)){
    if(is.interactive){report.progress(i, length(common.transcripts))}
    trans.id <- common.transcripts[i]
    trans.idx <- which(expr.annot[,1] == trans.id)
    trans.chr <- expr.annot[trans.idx,"chr"]
    trans.start <- expr.annot[trans.idx,"start"]
    trans.end <- expr.annot[trans.idx,"end"]
    id.locale <- which(expr.lod$additive[,1] == trans.id)
    on.chr <- which(lod.marker.chr == trans.chr)
    above.min <- which(lod.marker.pos >= (trans.start*1e6-bp.buffer))
    below.max <- which(lod.marker.pos <= (trans.end*1e6+bp.buffer))
    cis.locale <- Reduce("intersect", list(id.locale, on.chr, above.min, below.max))
    if(length(cis.locale) == 1){
        cis.lod[i] <- expr.lod$additive[cis.locale,"lod"]
    }
    if(length(cis.locale) > 1){stop}
}

plot.with.model(cis.lod, all.cor, xlab = "Local LOD Score", 
    ylab = "Correlation between imputed and measured")
```

## Diplotypes

It is possible that dominance, recessiveness, and epistasis 
could be affecting the eQTLs in ways that we are not measuring.

Our first step will be to look at DO diplotypes. What is the
distribution of diplotypes at any given locus in the DO? What
about the corresponding diplotypes in the CC-RIX? 

What is the mean expression for each diplotype in the DO? CC-RIX?
Do these match better than the individual haplotypes?

```{r diplotypes}

#change this to get cis marker
get_cis_marker <- function(transcript.id, gene.table, pop.genoprobs, pop.map, 
    round.probs = TRUE){

    trans.idx <- which(gene.table[,1] == transcript.id)[1]
    #gene.table[trans.idx,]
    trans.chr <- gene.table[trans.idx,"chr"]
    trans.start <- as.numeric(gene.table[trans.idx,"start"])
    cis.marker <- find_marker(pop.map, trans.chr, pos = trans.start)
    cis.geno <- pop.genoprobs[[trans.chr]][,,cis.marker]
    if(round.probs){
        cis.geno <- apply(cis.geno, 2, bin.vector)
    }
    return(cis.geno)
}

diplotype_idx <- function(marker.geno, present.thresh = 0.3){
    hap.present <- apply(marker.geno, 1, function(x) which(x > present.thresh))
    allele.pairs <- pair.matrix(colnames(marker.geno), self.pairs = TRUE)
    dip.idx <- vector(mode = "list", length = nrow(allele.pairs))
    names(dip.idx) <- apply(allele.pairs, 1, function(x) paste(x, collapse = "_"))
    for(i in 1:length(hap.present)){
        haps <- names(hap.present[[i]])
        if(length(haps) == 1){haps <- c(haps, haps)}
        test1 <- intersect(which(allele.pairs[,1] == haps[1]), which(allele.pairs[,2] == haps[2]))
        test2 <- intersect(which(allele.pairs[,2] == haps[1]), which(allele.pairs[,1] == haps[2]))
        hap.idx <- unique(c(test1, test2))
        dip.idx[[hap.idx]] <- c(dip.idx[[hap.idx]], i)
    }
    return(dip.idx)
}

dip.vals <- function(marker.geno, dip.idx, pheno.vals, cluster.by.val = TRUE){
    val.mat <- matrix(0, ncol = ncol(marker.geno), nrow = ncol(marker.geno))
    rownames(val.mat) <- colnames(val.mat) <- colnames(marker.geno)
    for(i in 1:length(dip.idx)){
        dip.type <- strsplit(names(dip.idx)[i], "_")[[1]]
        val.mat[dip.type[1], dip.type[2]] <- median(pheno.vals[dip.idx[[i]]])
        val.mat[dip.type[2], dip.type[1]] <- median(pheno.vals[dip.idx[[i]]])
    }
    if(cluster.by.val){
        row.order <- hclust(dist(val.mat))$order
    }else{
        row.order <- 1:8
    }
    imageWithText(val.mat[row.order, row.order], use.pheatmap.colors = TRUE, col.text.rotation = 0,
    col.text.shift = 0.1, row.text.shift = 0.1, show.text = FALSE, 
    main = "Median Expression by Diplotype")
    return(val.mat)
}

#look for dominance of individual alleles
allele_effect_dist <- function(marker.geno, pheno.vals){
    genotype.vals <- c(0, 0.5, 1)
    allele.dist <- apply(marker.geno, 2, 
        function(x) sapply(genotype.vals, function(y) pheno.vals[which(x == y)]))
    new.list <- lapply(1:length(allele.dist[[1]]), function(x) lapply(allele.dist, function(y) y[[x]]))
    names(new.list) <- genotype.vals
    plot.grouped.boxes(new.list, print.vals = NA)
    #tests <- lapply(allele.dist, aov_list)
    #p <- sapply(tests, function(x) summary(x)[[1]]$"Pr(>F)"[1])
}

#create a diploid genotype matrix
dip_geno <- function(dip.idx){
    dip.mat <- matrix(0, ncol = length(dip.idx), nrow = length(unlist(dip.idx)))
    colnames(dip.mat) <- names(dip.idx)
    for(i in 1:length(dip.idx)){
        dip.mat[dip.idx[[i]],i] <- 1
    }
    return(dip.mat)
}

add_pred <- function(allele.coef){
    pair.mat <- pair.matrix(allele.coef, self.pairs = TRUE)
    add.effect <- rowMeans(pair.mat)
    return(add.effect)
}

plot_dev <- function(transcript.id, gene.table, pop.genoprobs, pop.map, pop.expr,
    bin.genoprobs = TRUE, cluster.med.effect = FALSE, add.predict = NULL, png.name = NULL){

    cis.geno <- get_cis_marker(transcript.id, gene.table, pop.genoprobs, pop.map, bin.genoprobs)
    matched.geno <- get.xz(cis.geno, pop.expr)
    dip.idx <- diplotype_idx(marker.geno = matched.geno[[1]])

    if(!is.null(png.name)){
        png(png.name, width = 7, height = 10, units = "in", res = 180)
    }

    layout.matrix  <- matrix(c(1,5,2,2,3,3,4,4), ncol = 2, byrow = TRUE)
    layout(layout.matrix)

    par(mar = c(1,1,2,1))
    val.mat <- dip.vals(matched.geno[[1]], dip.idx, matched.geno[[2]][,transcript.id], 
        cluster.by.val = cluster.med.effect)    

    par(mar = c(4,4,2,2))
    allele_effect_dist(marker.geno = matched.geno[[1]], 
        pheno.vals = matched.geno[[2]][,transcript.id])    

    marker.scan <- fit1(matched.geno[[1]], matched.geno[[2]][,transcript.id])
    #quartz();barplot(marker.scan$coef[1:8])
    #plot(matched.geno[[2]][,transcript.id]~matched.geno[[1]][,"G"])
    #model <- lm(matched.geno[[2]][,transcript.id]~matched.geno[[1]][,"G"])
    if(is.null(add.predict)){
        dip.predicted <- add_pred(marker.scan$coef[1:8])
    }else{
        dip.predicted <- add.predict
    }

    predicted.order <- order(dip.predicted)

    dip.expr <- lapply(dip.idx, function(x) matched.geno[[2]][x,transcript.id])
    med.centered <- lapply(dip.expr, function(x) x - median(unlist(dip.expr)))

    ylim = c(min(c(unlist(med.centered), dip.predicted)), max(c(unlist(med.centered), dip.predicted)))

    boxplot(med.centered[predicted.order], las = 2, ylim = ylim)
    #stripchart(med.centered[predicted.order], add = TRUE, vertical = TRUE, pch = 16, col = "darkgray", cex = 0.5)
    segments(x0 = 0, x1 = length(med.centered)+1, y0 = 0)
    points(x = 1:length(dip.predicted), y = dip.predicted[predicted.order], col = "red", pch = "*", cex = 2)
    deviation <- sapply(med.centered, median) - dip.predicted
    barplot(deviation[predicted.order], las = 2, ylab = "Deviation from Additivity")

    dev.mat <- matrix(NA, nrow = ncol(cis.geno), ncol = ncol(cis.geno))
    colnames(dev.mat) <- rownames(dev.mat) <- colnames(cis.geno)
    for(i in 1:length(deviation)){
        dip.type <- strsplit(names(deviation)[i], "_")[[1]]
        dev.mat[dip.type[1], dip.type[2]] <- deviation[i]
        dev.mat[dip.type[2], dip.type[1]] <- deviation[i]
    }
    
    par(mar = c(1,1,2,1))
    imageWithText(dev.mat, use.pheatmap.colors = TRUE, col.text.rotation = 0,
    col.text.shift = 0.1, row.text.shift = 0.1, show.text = FALSE, 
    main = "Deviation by Diplotype")

    #plot.with.model(as.vector(val.mat), as.vector(dev.mat))

    #hist(deviation)

    if(!is.null(png.name)){
        dev.off()
    }

    result <- cbind(dip.predicted, deviation)
    colnames(result)  <- c("Additive_Prediction", "Deviation")
    return(result)

}

#calculates values similarly to plot_dev, but without plotting
calc_dev <- function(transcript.id, gene.table, pop.genoprobs, pop.map, pop.expr,
    bin.genoprobs = FALSE, add.predict = NULL, plot.results = FALSE,
    present.thresh = 0.8){
    
    cis.geno <- get_cis_marker(transcript.id, gene.table, pop.genoprobs, pop.map, bin.genoprobs)
    matched.geno <- get.xz(cis.geno, pop.expr)
    marker.scan <- fit1(matched.geno[[1]], matched.geno[[2]][,transcript.id])
    #barplot(marker.scan$coef[1:8])    
    mean.expr <- mean(matched.geno[[2]][,transcript.id])

    if(is.null(add.predict)){
        dip.predicted <- add_pred(marker.scan$coef[1:8])
    }else{
        dip.predicted <- add.predict
    }

    allele.pairs <- pair.matrix(1:ncol(matched.geno[[1]]), self.pairs = TRUE)
    actual.effect <- rep(NA, length = nrow(allele.pairs))
    for(i in 1:nrow(allele.pairs)){
        dip.genotype <- rowMeans(matched.geno[[1]][,allele.pairs[i,]])
        dip.locale <- which(dip.genotype > present.thresh)
        #model <- lm(matched.geno[[2]][,transcript.id]~dip.genotype)
        #plot.with.model(dip.genotype, matched.geno[[2]][,transcript.id], report = "cor.test")
        actual.effect[i] <- mean.expr - mean(matched.geno[[2]][dip.locale,transcript.id])
    }

    pair.labels <- apply(allele.pairs, 1, function(x) paste(LETTERS[x], collapse = "_"))

    if(plot.results){
        plot(dip.predicted, actual.effect, xlab = "Additive Prediction", 
            ylab = "Actual Effect", pch = 16, cex = 0.5)
        abline(0,1)
        text(dip.predicted, actual.effect, labels = pair.labels)
    }
    
    add.dev <- dip.predicted - pred.expr
    names(add.dev) <- pair.labels
    result <- cbind(pred.expr, dip.predicted)
    colnames(result) <- c("Additive_Prediction", "Actual_Expression")
    rownames(result) <- pair.labels
    return(result)
}


transcript.id <- sample(common.transcripts, 1)
transcript.id <- names(cis.lod)[which(cis.lod > 200)][1]
transcript.id <- "ENSMUSG00000028619" #transcript for which E/E is very low

#do dev
do.dev <- plot_dev(transcript.id, gene.table, pop.genoprobs = genoprobs, 
    pop.map = map, pop.expr = do.adj.expr, bin.genoprobs = TRUE, 
    cluster.med.effect = FALSE, add.predict = NULL, 
    png.name = here("Results", "CCA_Clusters", "Validation", "Deviation.DO.png"))

#test.dev <- calc_dev(transcript.id, gene.table, pop.genoprobs = genoprobs, 
#    pop.map = map, pop.expr = do.adj.expr, bin.genoprobs = FALSE, add.predict = NULL,
#    plot.results = TRUE)

#plot(do.dev)
#plot(test.dev);abline(0,1)
#plot(do.dev[,1], test.dev[,1])
#plot(do.dev[,2], test.dev[,2])

#CC-RIX deviation relative to DO predictions
cc.dev <- plot_dev(transcript.id, gene.table, pop.genoprobs = expanded.geno, 
    pop.map = cc.map, pop.expr = adj.cc.expr, bin.genoprobs = TRUE, 
    cluster.med.effect = FALSE, add.predict = do.dev[,"Additive_Prediction"], 
    png.name = here("Results", "CCA_Clusters", "Validation", "Deviation.CC.png"))

plot.with.model(do.dev[,2], cc.dev[,2], xlab = "DO deviation from additivity",
    ylab = "CC deviation from DO additivity")


#dip.geno <- dip_geno(dip.idx)
#rownames(dip.geno) <- rownames(matched.do.geno[[1]])
#dip.scan <- fit1(dip.geno, do.adj.expr[,transcript.id])
#quartz();barplot(sort(dip.scan$coef[1:length(dip.scan$coef)-1]), las = 2)
#allele_effect_dist(matched.do.geno[[1]], matched.do.geno[[2]][,transcript.id])
#marker.geno <- matched.do.geno[[1]]
#pheno.vals <- matched.do.geno[[2]][,transcript.id]

```

## Example comparisons

```{r var}
#This code spot checks genes to compare within-strain variance
#and between-strain variance across the CC with the imputed 
#values and local LOD scores.

plot_cc_v_imp <- function(transcript.id, gene.table, pop.genoprobs, pop.map, bin.genoprobs){

    cis.geno <- get_cis_marker(transcript.id, gene.table, pop.genoprobs, pop.map, bin.genoprobs)
    cross.geno <- sapply(cross_expr, function(x) cis.geno[which(rownames(cis.geno) %in% rownames(x))[1],])

    test.expr <- lapply(cross_expr, function(x) x[,transcript.id])
    mean.test <- sapply(test.expr, mean)

    mean.order <- order(mean.test)
    layout.mat <- matrix(c(1,2,3,0,4,0), byrow = FALSE, ncol = 2)
    
    layout(layout.mat, widths = c(1, 0.3))
    par(mar = c(0, 4, 2, 4))
    local.lod <- round(cis.lod[transcript.id])
    if(is.na(local.lod)){local.lod <- "< 6"}
    barplot(imp.auto.expr[mean.order,transcript.id], las = 2, 
        main = paste0(transcript.id, ", LOD:" , local.lod),
        axes = FALSE, names = NA)
    mtext("Imputed Expression", side = 2, line = 2.5)
    axis(2)
    par(mar = c(0,4,0,4))
    boxplot(test.expr[mean.order], names = NA, las = 2, axes = FALSE)
    mtext("Measured Expression", side = 2, line = 2.5)
    axis(2)
    plot.dim <- par("usr")
    
    par(mar = c(8,4,0,4))
    imageWithText(cross.geno[,mean.order], use.pheatmap.colors = TRUE, 
        row.text.shift = 0.05)

    instrain.var <- sapply(test.expr, var)
    betstrain.var <- var(sapply(test.expr, mean))
    xlim <- c(min(c(instrain.var, betstrain.var)), max(c(instrain.var, betstrain.var)))
    par(mar = c(0,0,2,3))
    hist(instrain.var, xlim = xlim, main = "Within-Strain vs.\nBetween-Strain Variation")
    abline(v = betstrain.var, col = "red")
}

```

There is no relationship between the imputed CC-RIX expression values
and the measured values. I wanted to dig into this a bit more to test
whether this is a bug somewhere along the line or whether we really 
can't impute CC-RIX gene expression from DO haplotype effects.

For some transcripts in the CC-RIX there are large across-strain
differences. For example...

```{r hi_strain_var}

test.idx <- which(cis.lod > 200)
idx <- 1
transcript.id <- names(test.idx)[idx]

plot_cc_v_imp(transcript.id, gene.table, pop.genoprobs = expanded.geno, 
    pop.map = cc.map, bin.genoprobs = TRUE)
```

```{r, low_strain_var}
test.idx <- which(cis.lod < 6.5)
idx <- 1
transcript.id <- names(test.idx)[idx]

plot_cc_v_imp(transcript.id, gene.table, pop.genoprobs = expanded.geno, 
    pop.map = cc.map, bin.genoprobs = TRUE)
```

## Inter-strain variation

If there is a relationship between LOD score in the DO and 
across strain variation in the CC, we would expect to see
greater between-strain variation relative to within-strain
variation in transcripts with higher LOD scores in the DO. 

We used ANOVA to calculate an F statistic for each transcript.
The larget the F statistic, the more different the strains are 
in expression.

There does actually appear to be a relationship between LOD 
score in the DO, and across-strain variation in the CC-RIX.

```{r var_vs_lod}
get_f <- function(transcript.id, strain.factor){
    test.expr <- lapply(cross_expr, function(x) x[,transcript.id])
    model <- aov(unlist(test.expr)~strain.factor)
    f <- summary(model)[[1]]$"F value"[1]
    return(f)
}

f.file <- here("Results", "CCA_Clusters", "Validation", "CC-RIX_transcript_F_tests.RDS")
if(!file.exists(f.file)){
    strain.list <- lapply(1:length(cross_expr), function(x) rep(names(cross_expr)[x], nrow(cross_expr[[x]])))
    strain.factor <- as.factor(unlist(strain.list))
    f <- sapply(common.transcripts, function(x) get_f(x, strain.factor))
    names(f) <- common.transcripts
    saveRDS(f, f.file)
}else{
    f <- readRDS(f.file)
}

plot.with.model(log(cis.lod), log(f), xlab = "log(Local LOD score)", 
    ylab = "Across-Strain Variance (log(F))", report = "cor.test")

#test.idx <- which(f > 100)
#test.idx <- which(f < 1)
#plot_cc_v_imp(names(test.idx)[1])
```

## LOD score and variance components

DO cis eQTL LOD score should correlate very highly with the
amount of variance explained by the local eQTL.

```{r lod_local_var}
plot.with.model(cis.lod, var.comp[common.transcripts,"local_VE"],
    xlab = "Local LOD score", ylab = "Variance Explained by Local eQTL")
```

There are postive correlations between across-strain variation in 
the CC-RIX (F), local LOD scores in the DO, and variance explained
by the local genetics.


```{r local_var_f}
comp.table <- cbind(log(f), var.comp[common.transcripts,], cis.lod)
colnames(comp.table) <- c("log(F)", colnames(var.comp), "Local LOD")
pairs(comp.table)
```

## Local VE and imputed-measured correlation

The following plot shows how the variance explained by 
local genetics in the DO relates to the correlation between
measured and imputed expression in the CC-RIX. It doesn't.

```{r lod_cor_ve}
plot.with.model(var.comp[common.transcripts,"local_VE"], all.cor,
xlab = "VE local", ylab = "Correlation between measured and imputed transcripts in CC-RIX")
```

## Spot Checking

We are going to look at individual genes with strange combinations
of values from above to see if we can dig deeper into what is going
on.


The following histogram shows the overall distribution of correlations
between DO allele coefficients and CC coefficients at the marker
for the cis eQTL in the DO.

```{r, do_cc_cor}

#to parallelize the DO and CC analyses, use fit1 for 
#both
get_transcript_pos <- function(transcript.id){
    trans.idx <- which(expr.annot[,1] == transcript.id)
    trans.chr <- expr.annot[trans.idx,"chr"]
    trans.start <- expr.annot[trans.idx,"start"]
    return(c(trans.chr, trans.start))
}

get_coef <- function(transcript.chr, transcript.pos, mouse.expr, mouse.geno, mouse.map, 
    mouse.K, mouse.covar, use.kinship = TRUE, plot.results = FALSE){
    
    cis.marker <- find_marker(mouse.map, transcript.chr, transcript.pos)
    
    #bin the genotype into 0, 0.5, and 1. Otherwise we get super crazy coefficients
    binned.geno <- apply(mouse.geno[[transcript.chr]][,,cis.marker], 2, bin.vector)
    if(use.kinship){
        norm.expr <- mouse.expr[,transcript.id,drop=FALSE]
        names(norm.expr) <- rownames(mouse.expr)
        marker.fit <- fit1(binned.geno, pheno = norm.expr, 
            kinship = mouse.K[[transcript.chr]],
            addcovar = mouse.covar)
    }else{
        norm.expr <- mouse.expr[,transcript.id,drop=FALSE]
        names(norm.expr) <- rownames(mouse.expr)
        marker.fit <- fit1(binned.geno, pheno = norm.expr, 
            addcovar = mouse.covar)
    }
    
    mouse.coef <- marker.fit$coef[1:8]

    if(plot.results){
        coef.order <- order(mouse.coef)
        barplot(mouse.coef[coef.order], col = CCcolors[coef.order])
    }

    return(mouse.coef)
}

#This function does not use fit1, but fits a linear model to 
#each allele separately. This method does take covariates into 
#account, but not kinship. This method also 

get_coef_by_allele <- function(transcript.chr, transcript.pos, mouse.expr, 
    mouse.geno, mouse.map, mouse.K, mouse.covar, plot.results = FALSE){
    
    cis.marker <- find_marker(mouse.map, transcript.chr, transcript.pos)
    binned.geno <- apply(mouse.geno[[transcript.chr]][,,cis.marker], 2, bin.vector)
    common.ind <- Reduce("intersect", list(rownames(binned.geno), rownames(mouse.expr), rownames(mouse.covar)))
    common.geno.ind <- match(common.ind, rownames(binned.geno))
    common.expr.ind <- match(common.ind, rownames(mouse.expr))
    common.covar.ind <- match(common.ind, rownames(mouse.covar))
    allele.tests <- apply(binned.geno[common.geno.ind,], 2, 
        function(x) lm(rankZ(mouse.expr[common.expr.ind,transcript.id,drop=FALSE])~mouse.covar[common.covar.ind,]+x))
    allele.coef <- t(sapply(allele.tests, function(x) c(head(coef(x), 1), tail(coef(x), 1))))
    mouse.coef <- allele.coef[,2] - allele.coef[,1]
    #barplot(allele.coef[,2])
    #barplot(mouse.coef)
    #plot(allele.coef[,2], mouse.coef, col = CCcolors, pch = 16, cex = 2)

    marker.fit <- fit1(binned.geno, 
        pheno = mouse.expr[,transcript.id,drop=FALSE], kinship = mouse.K[[transcript.chr]],
        addcovar = mouse.covar)
    
    mouse.coef <- marker.fit$coef[1:8]

    if(plot.results){
        par(mfrow = c(1,2))
        plot.with.model(allele.coef[,2], mouse.coef, col = CCcolors, pch = 16, cex = 2, 
            main = "Allele Effects", xlab = "fit1 with kinship",
            ylab = "allele lm")
        plot.with.model(rankZ(allele.coef[,2]), rankZ(mouse.coef), col = CCcolors, 
            pch = 16, cex = 2, main = "RankZ Allele Effects",
            xlab = "rankZ fit1 with kinship coef", ylab = "rankZ allele lm")
        #coef.order <- order(mouse.coef)
        #barplot(rankZ(mouse.coef[coef.order]), col = CCcolors[coef.order])
    }
    result <- rbind(allele.coef[,2], mouse.coef)
    rownames(result) <- c("individual_fit", "fit_with_kinship")
    return(result)
}

get_num_trans <- function(transcript.id, bp.buffer = 4e6){
    transcript.idx <- which(do.expr$lod.peaks$additive[,1] == transcript.id)
    if(length(transcript.idx) == 0){return(0)}
    #do.expr$lod.peaks$additive[transcript.idx,]
    transcript.pos <- get_transcript_pos(transcript.id)
    qtl.markers <- do.expr$lod.peaks$additive[transcript.idx,2]
    split.markers <- strsplit(qtl.markers, "_")
    qtl.chr <- sapply(split.markers, function(x) x[1])
    qtl.pos <- sapply(split.markers, function(x) as.numeric(x[2]))
    chr.locale <- which(qtl.chr == transcript.pos[1])
    above.min <- which(qtl.pos/1e6 >= (as.numeric(transcript.pos[2])-(bp.buffer/1e6)))
    below.max <- which(qtl.pos/1e6 <= (as.numeric(transcript.pos[2])+(bp.buffer/1e6)))
    cis.idx <- Reduce("intersect", list(chr.locale, above.min, below.max))
    trans.idx <- setdiff(transcript.idx, cis.idx)
    return(length(trans.idx))
}

#transcript.id <- sample(common.transcripts, 1) #random transcript
plot_do_cc_coef <- function(transcript.id, bp.buffer = 4e6, 
    cor.type = c("Spearman", "Pearson"), do.kinship = TRUE, 
    cc.kinship = FALSE, plot.results = TRUE){

        cor.type = cor.type[1]
        transcript.info <- get_transcript_pos(transcript.id)
        transcript.chr <- transcript.info[1]
        transcript.pos <- as.numeric(transcript.info[2])

        #get allele coefficients using fit1 and a kinship correction
        do.coef <- get_coef(transcript.chr, transcript.pos, mouse.expr = do.expr$data$norm, 
            mouse.geno = genoprobs, mouse.map = map, mouse.K = K, 
            mouse.covar = do.expr$covar.matrix, use.kinship = do.kinship, 
            plot.results = FALSE)

        if(is.interactive){
            #compare allele coefficients using fit1 and independent model fitting
            quartz(width = 8, height = 4)
            test.do.coef <- get_coef_by_allele(transcript.chr, transcript.pos, 
                mouse.expr = do.expr$data$norm, 
                mouse.geno = genoprobs, mouse.map = map, mouse.K = K, 
                mouse.covar = do.expr$covar.matrix, plot.results = TRUE)
                #plot(do.coef, test.do.coef[2,], col = CCcolors, pch = 16, cex = 2);abline(0,1)
            mtext("DO", side = 3, outer = TRUE, line = -1.5, font = 2)
        }

        cc.coef <- get_coef(transcript.chr, transcript.pos, 
            mouse.expr = tissue.expr[[tissue.locale]], 
            mouse.geno = expanded.geno, mouse.map = cc.map, mouse.K = expanded.K, 
            mouse.covar = rnaseq.covar, plot.results = FALSE, use.kinship = cc.kinship)

        if(is.interactive){
            quartz(width = 8, height = 4)
            test.cc.coef <- get_coef_by_allele(transcript.chr, transcript.pos, 
                mouse.expr = tissue.expr[[tissue.locale]], 
                mouse.geno = expanded.geno, mouse.map = cc.map, mouse.K = expanded.K, 
                mouse.covar = rnaseq.covar, plot.results = TRUE)
            mtext("CC-RIX", side = 3, outer = TRUE, line = -1.5, font = 2)
            #plot(cc.coef, test.cc.coef[2,], col = CCcolors, pch = 16, cex = 2);abline(0,1)
        }

        if(is.interactive){
            #compare DO and CC effects to each other with and without kinship
            all.coef <- cbind(t(test.do.coef), t(test.cc.coef))
            colnames(all.coef) <- paste0(c(rep("DO_", 2), rep("CC-RIX_", 2)), colnames(all.coef))
            type.pairs <- pair.matrix(1:4)
            quartz()
            par(mfrow = c(2,2))
            for(i in 2:(nrow(type.pairs)-1)){
                type1 <- type.pairs[i,1]
                type2 <- type.pairs[i,2]
                plot.with.model(all.coef[,type1], all.coef[, type2], 
                xlab = colnames(all.coef)[type1], ylab = colnames(all.coef)[type2], 
                col = CCcolors, cex = 2, report = "cor.test", 
                main = paste(colnames(all.coef)[type1], "vs.\n", colnames(all.coef)[type2]))

            }

        }

        num.trans <- get_num_trans(transcript.id, bp.buffer)

        if(cor.type == "Spearman"){
            norm.do.coef <- rankZ(do.coef)
            norm.cc.coef <- rankZ(cc.coef)
        }else{
            norm.do.coef <- do.coef
            norm.cc.coef <- cc.coef
        }

        if(plot.results){
            if(is.interactive){quartz()}
            r <- plot.with.model(norm.do.coef, norm.cc.coef,
                xlab = "DO Coefficient", ylab = "CC Coefficient",
                pch = 16, col = CCcolors, cex = 2, report = "cor.test",
                main = transcript.id)[1]
        }else{
            r <- cor(norm.do.coef, norm.cc.coef, 
                use = "pairwise.complete.obs")
        }

    result <- c(r, num.trans)
    names(result) <- c("r", "num_trans")
    invisible(result)
}

if(is.interactive){

    plot.with.model(var.comp[common.transcripts,"local_VE"], all.cor,
    xlab = "VE local", ylab = "Correlation between measured and imputed transcripts in CC-RIX")

    ve.local <- var.comp[common.transcripts,"local_VE"]

    test.transcript <- common.transcripts[intersect(which(ve.local > 0.7), which(all.cor < -0.5))]
    test.transcript <- common.transcripts[intersect(which(ve.local > 0.7), which(all.cor > 0.6))]
    
    rnd.transcript <- sample(test.transcript, 1)
    #rnd.transcript <- test.transcript[1]
    #rnd.transcript <- transcript.id

    tr.pos <- get_transcript_pos(rnd.transcript)
    do.cis.marker <- find_marker(map, tr.pos[1], as.numeric(tr.pos[2]))
    do.cis.geno <- genoprobs[[tr.pos[1]]][,,do.cis.marker]
    do.freq <- colMeans(do.cis.geno)  

    cc.cis.marker <- find_marker(cc.map, tr.pos[1], as.numeric(tr.pos[2]))
    cc.cis.geno <- expanded.geno[[tr.pos[1]]][,,cc.cis.marker]
    cc.freq <- colMeans(cc.cis.geno)
    
    #quartz();barplot(rbind(do.freq,cc.freq), beside = TRUE, col = rep(CCcolors, each = 2))

    plot_do_cc_coef(rnd.transcript, plot.results = TRUE, 
        cor.type = "Spearman", do.kinship = TRUE, cc.kinship = TRUE)
    
    plot_do_cc_coef(rnd.transcript, plot.results = TRUE, 
        cor.type = "Pearson", do.kinship = TRUE, cc.kinship = TRUE)
}

do.cc.cor.kin.file <- here("Results", "CCA_Clusters", "imputed_local", paste0("cc.do.cor.", tissue.name, ".kin.RDS"))
do.cc.cor.file <- here("Results", "CCA_Clusters", "imputed_locale", paste0("cc.do.cor.", tissue.name, ".RDS"))
if(!file.exists(do.cc.cor.file)){
    all.do.coef.cor.kin <- t(sapply(names(cis.lod), 
        function(x) plot_do_cc_coef(x, plot.results = FALSE, 
        cor.type = "Spearman", do.kinship = TRUE, cc.kinship = TRUE)))
    saveRDS(all.do.coef.cor.kin, do.cc.cor.kin.file)
    all.do.coef.cor <- t(sapply(names(cis.lod), 
        function(x) plot_do_cc_coef(x, plot.results = FALSE, 
        cor.type = "Spearman", do.kinship = TRUE, cc.kinship = FALSE)))
    saveRDS(all.do.coef.cor, do.cc.cor.file)
}else{
    all.do.coef.cor.kin <- readRDS(do.cc.cor.kin.file)
    all.do.coef.cor <- readRDS(do.cc.cor.file)
}


hist(all.do.coef.cor[test.idx,1], breaks = 100, 
    main = "Without Kinship",
    xlab = "Spearman Rank Correlation")
abline(v = median(all.do.coef.cor[test.idx,1], na.rm = TRUE), col = "red")

hist(all.do.coef.cor.kin[test.idx,1], breaks = 100, 
    main = "With Kinship",
    xlab = "Spearman Rank Correlation")
abline(v = median(all.do.coef.cor.kin[test.idx,1], na.rm = TRUE), col = "red")

```


The following plot shows how the correlation between allele effects
in the DO and CC depend on LOD score in the DO. We expect a priori
that transcripts with high LOD scores in the DO will have a higher
correlation of allele coefficients than transcripts with low LOD
scores.

```{r lod_cor}
plot.with.model(cis.lod, all.do.coef.cor[,1], report = "cor.test", 
    xlab = "DO Local LOD score",
    ylab = "Correlation between DO coefficients and CC coefficients")
```

The plot below shows that the correlation between DO and CC-RIX
coefficients is not related to the number of trans eQTL in the DO.

```{r trans_check}


plot.with.model(all.do.coef.cor[,1], all.do.coef.cor[,2], report = "cor.test", 
    ylab = "Number of Trans eQTL",
    xlab = "Correlation between DO coefficients and CC coefficients",
    main = "without kinship")

plot.with.model(all.do.coef.cor.kin[,1], all.do.coef.cor.kin[,2], report = "cor.test", 
    ylab = "Number of Trans eQTL",
    xlab = "Correlation between DO coefficients and CC coefficients",
    main = "with kinship")


#high.cor <- which(all.do.coef.cor[,1] > 0.9)
#plot_do_cc_coef(names(high.cor)[6], plot.results = TRUE)
```

```{r compare_do_cc}
compare_do_cc <- function(transcript.id){
    
    transcript.info <- get_transcript_pos(transcript.id)
    transcript.chr <- transcript.info[1]
    transcript.pos <- as.numeric(transcript.info[2])

    do.coef <- get_coef(transcript.chr, transcript.pos, mouse.expr = do.expr$data$rz, 
        mouse.geno = genoprobs, mouse.map = map, mouse.K = K, 
        mouse.covar = do.expr$covar.matrix, plot.results = FALSE)

    cc.coef <- get_coef(transcript.chr, transcript.pos, 
        mouse.expr = tissue.expr[[tissue.locale]], 
        mouse.geno = expanded.geno, mouse.map = cc.map, mouse.K = expanded.K, 
        mouse.covar = rnaseq.covar, plot.results = FALSE)


    do.cis.marker <- find_marker(map, trans.chr, transcript.pos)
    binned.do.geno <- apply(genoprobs[[trans.chr]][,,do.cis.marker],2, bin.vector)
    adj.do.expr <- adjust(do.expr$data$norm[,transcript.id,drop=FALSE], do.expr$covar.matrix)
    common.do.ind <- intersect(rownames(adj.do.expr), rownames(binned.do.geno))
    common.do.trans.idx <- match(common.do.ind, rownames(adj.do.expr))
    common.do.geno.idx <- match(common.do.ind, rownames(binned.do.geno))

    cc.cis.marker <- find_marker(cc.map, trans.chr, transcript.pos)
    binned.cc.geno <- apply(expanded.geno[[trans.chr]][,,cc.cis.marker],2, bin.vector)
    adj.cc.expr <- adjust(cc.expr[,transcript.id,drop=FALSE], rnaseq.covar)
    common.cc.ind <- intersect(rownames(adj.cc.expr), rownames(binned.cc.geno))
    common.cc.trans.idx <- match(common.cc.ind, rownames(adj.cc.expr))
    common.cc.geno.idx <- match(common.cc.ind, rownames(binned.cc.geno))

    par(mfrow = c(4,4), mar = c(1,1,1,1))    
    #barplot(unlist(expr.lod$additive[cis.locale,LETTERS[1:8][do.coef.order]]))
    for(cc in 1:8){
        boxplot(adj.do.expr[common.do.trans.idx]~binned.do.geno[common.do.geno.idx,cc], 
        col = CCcolors[cc], xlab = "", ylab = "DO expression", 
        main = paste("DO", names(CCcolors)[cc]), axes = FALSE)
        axis(2)
    

        boxplot(adj.cc.expr[common.cc.trans.idx]~binned.cc.geno[common.cc.geno.idx,cc], 
        col = CCcolors[cc], xlab = "", ylab = "CC expression", 
        main = paste("CC", names(CCcolors)[cc]), axes = FALSE)
        axis(2)
    }
}

```

```{r spot_check, eval = FALSE}
check.idx <- intersect(which(all.do.coef.cor[,1] < -0.5), which(cis.lod > 100))

check.idx <- intersect(which(all.do.coef.cor[,1] > 0.9), which(cis.lod > 100))
check.idx <- which(cis.lod > 150)

check.id <- names(cis.lod)[check.idx[6]]
check.id <- names(cis.lod)[2]
check.id <- "ENSMUSG00000031101" #Sash3
check.id <- "ENSMUSG00000028619"


quartz();plot_do_cc_coef(check.id, cor.type = "Spearman", do.kinship = TRUE, cc.kinship = FALSE)
quartz();plot_do_cc_coef(check.id, cor.type = "Spearman", do.kinship = TRUE, cc.kinship = TRUE)
quartz();plot_do_cc_coef(check.id, cor.type = "Spearman", do.kinship = FALSE, cc.kinship = FALSE)

quartz();compare_do_cc(check.id)


```