---
title: "High Dimensional Mediation TRiAGE"
author: Anna L Tyler
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    collapsed: no
    toc: yes
    toc_float: yes
---

## Introduction
This workflow explores the possibility of using TRiAGE to interpret
high-dimensional mediation results.

```{r code}
library(here)
all.fun <- list.files(here("Code"), pattern = ".R", full.names = TRUE)
for(i in 1:length(all.fun)){source(all.fun[i])}
```

```{r libraries}
needed.libraries <- c("Matrix", "knitr", "here", "igraph", "DescTools", "pheatmap", 
"e1071", "parallel", "doParallel", "gprofiler2", "clusterProfiler", "DT",
"plotly", "qtl2") 
load_libraries(needed.libraries)
```

```{r param}
fntm.tissue.name <- "adipose tissue"; tissue.name <- "Adipose"
n.genes = 250 #number of genes to select from top and bottom of the list
results.dir <-  here("Results", "Prioritization")
project.name <- paste("HDM", tissue.name, sep = "_")
n.cores = detectCores()-1
```

```{r read_data}
#exp.name <- "tissue_sep-full_imp-_-complete_mediation"
exp.name <- "tissue_together-local_imp-_-complete_mediation"
mge <- readRDS(here("Results", "High_Dim_Med", exp.name, "Population_Pressure_Loadings.RDS"))
tissue.mge <- mge[[which(names(mge) == tissue.name)]]
triage.genes <- c(sort(tissue.mge, decreasing = TRUE)[1:n.genes], sort(tissue.mge, decreasing = FALSE)[1:n.genes])
mouse.gene.info <- read.delim(here("Data", "mouse_gene_info.txt"))
```


```{r download_net, eval = TRUE, echo = FALSE}
#Download the network if it hasn't been previously downloaded. 
#full list available at:  http://fntm.princeton.edu
tissue.type = gsub(" ", "_", fntm.tissue.name)
net.file <- here("Data", "FNTM", paste0(tissue.type, "_top.RDS"))
if(!file.exists(net.file)){
    download.tissue.net(tissue.type, organism = "mouse", top.edges.only = TRUE, 
    project.dir = here("Data"))
    }
```

```{r read_net, echo = FALSE}
#Read in the downloaded network.
all.var <- ls()
net.loaded <- as.logical(length(which(all.var == "tissue.net")))
if(!net.loaded){
    tissue.net <- readRDS(net.file)
}
```

Get entrez gene IDs for training set genes.

```{r entrez}
gene.id <- names(triage.genes)
entrezgene <- gconvert(gene.id, organism = "mmusculus", target = "ENTREZGENE_ACC")
```

## Setup Positive-Labeled Gene Features

Get the connection weights of the positive-labeled genes to each other
in the tissue network. Not all of the genes will have Entrez IDs, and 
not all of the genes will have connections to the other genes in the 
tissue network. We trim off positive-labeled genes that are not 
connected to any other positive-labeled genes.

```{r pos_net, echo = FALSE}
tissue.adj <- tissue.adj.mat(tissue.net, entrezgene[,"target"], inc.all.genes = FALSE)
```

After identifying all Entrezgene IDs and building the positive-labeled network,
we have `r nrow(tissue.adj)` positive-labeled genes for training.

## Train SVM
We used this positive-labeled network as features to train 100 SVMs to classify
TB- and not TB-related genes. In each of the 100 SVMs, we selected a balanced
set of unlabeled genes for training. We write each of the trained models to a 
file for further analysis.

```{r train_models, message = FALSE, warning = FALSE, error = FALSE, eval = TRUE}
generate.triage.models(path = results.dir, project.name = project.name, 
trait.genes = entrezgene[,"target"], fgn = tissue.net, n.trials = 100, 
cluster.modules = FALSE, cluster.threshold = 700, max.cluster.size = 600, 
min.cluster.size = 10, use.SVD = FALSE, verbose = TRUE, n.cores = n.cores)
```

## ROC Curves

We calculated ROC curves for each SVM, as well as the overal AUC for
the group. The plot below shows the ROC curves for all 100 models
along with the average AUC.

```{r roc, fig.height = 5, fig.width = 5}
all.fptp <- svm.ROC(results.dir, n.cores = n.cores, plot.results = TRUE)
all.auc <- sapply(all.fptp, function(x) smooth.auc(Reduce("rbind", x), plot = FALSE))
```

## Score all genes

We then used the trained models to classify all expressed genes.

```{r score_genes}
all.gene.id <- gconvert(names(tissue.mge), organism = "mmusculus", target = "ENTREZGENE_ACC")
score.candidates(results.dir, 
    candidate.genes = as.numeric(all.gene.id[,"target"]),
    verbose = TRUE, plot.results = FALSE)

merge.svm.gene.info(results.dir, mouse.gene.info)
```


```{r get_fp, error = FALSE, warning = FALSE, message = FALSE, fig.width = 8, fig.height = 5}
fp.mat <- get.all.fp(results.dir)
hist(-log10(fp.mat[,1]))

fp.order <- order(-log10(fp.mat[,1]), decreasing = TRUE)
#head(fp.mat[fp.order,,drop=FALSE])
```

There is a positive correlation between MGE and FPR.
Are genes that rank high on both scales good targets?
What about genes that rank highly on one or the other scale?

If the gene is high on just the TRiAGE scale, it might be a good target 
because it is highly related to the genes that are changing, but doesn't 
have a lot of genetic variation.

On the other hand, a gene that ranks high on both scales could 
be good, because we have a causal link between that gene and 
the phenotype.

```{r rank_genes}
fp.gene.id <- mouse.gene.info[match(rownames(fp.mat), mouse.gene.info[,"external_gene_name"]),"ensembl_gene_id"]
fp.idx <- match(fp.gene.id, names(tissue.mge))
plot.with.model(abs(tissue.mge[fp.idx]), -log10(fp.mat), xlab = "|MGE|",
    ylab = "-log10(FPR)")

ranked.genes <- rank.2D(cbind(abs(tissue.mge[fp.idx]), -log10(fp.mat)))
rank.order <- order(ranked.genes[,1], decreasing = FALSE)
head(ranked.genes[rank.order,,drop=FALSE])

rank.col <- colors.from.values(ranked.genes[,1], grad.dir = "low", col.scale = "blue")
plot.with.model(abs(tissue.mge[fp.idx]), -log10(fp.mat), xlab = "|MGE|",
    ylab = "-log10(FPR)", col = rank.col)
```

The following genes rank highly on both scales:

```{r both_scales}
mge.idx <- match(rownames(ranked.genes)[rank.order], names(tissue.mge))
fpr.idx <- match(rownames(ranked.genes)[rank.order], fp.gene.id)
merged.table <- data.frame("id" = rownames(ranked.genes)[rank.order], 
    "name" = rownames(fp.mat)[fpr.idx], 
    "rank" = ranked.genes[rank.order,], "MGE" = tissue.mge[mge.idx], 
    "neg_log10_FPR" = -log10(fp.mat[fpr.idx,1]))
```

Genes that are highly ranked on both axes are enriched for 
immune function. This is very common for these kinds of 
enrichment. It is definitely relevant, but I think immune
genes tend to vary more than other genes in general.

```{r top_enrich, fig.width = 7, fig.height = 5}
top.enrich <- gost(merged.table[1:50,1], organism = "mmusculus")
plot.enrichment(top.enrich, num.terms = 30)
```

The top-ranked gene for adipose tissues is Retsat. This gene has a large negative
MGE, meaning that decreased expression is associated with a more diabetes-like
phenotype. Retsat is retinol saturase and is involved in retinol metabolism.
It is located in the ER membrane and nuclear outer membrane. KO mice exhibit 
increased adiposity, consistent with the negative MGE.

What about genes that are high only on the TRiAGE scale?
These are enriched for metabolic processes. These are 
genes that either do not vary much, or whose variation
is not related to the traits, but are highly related
to the top MGE genes according to TRiAGE.

Where does the variance of these genes come from?

```{r var}
var.comp <- readRDS(here("Data", "imputed", paste0("DO_", tissue.name, "_imputed_variance_decomposition.RDS")))
var.idx <- match(merged.table[,"id"], rownames(var.comp))
merged.table <- cbind(merged.table, var.comp[var.idx,])

boxplot(merged.table[1:50,c("local_VE", "distal_VE")])
```

```{r triage_scale}
gene.idx <- order(merged.table[,"neg_log10_FPR"], decreasing = TRUE)[1:15]
just.triage <- merged.table[gene.idx,]
print(just.triage)

enrich <- gost(just.triage[,1], organism = "mmusculus")
plot.enrichment(enrich)
```

These genes have a greater distal component than the genes
that rank highly on both scales.

```{r triage_var}
triage.genes <- just.triage[,1] #for neg genes
triage.idx <-  match(triage.genes, merged.table[,"name"])
pairs(merged.table[triage.idx,c("local_VE", "distal_VE")])
```

The genes with a large local component just don't have any
effect on the trait, but the genes with a large distal 
component may be more relevant. The gene with the largest
distal component by far, is Anxa7. This gene is related
to insulin secretion in pancreatic beta cells, but doesn't
seem to have a known relationship in adipose tissue. Ugh.

```{r distal}
filtered.genes <- merged.table[which(merged.table[triage.idx,"distal_VE"] > 0.2),"name"]
```


Overall, the genes that are high on just the TRiAGE scal are enriched 
for mitochondrial function.

Hfe, Oas1g, and Bckdhb have all been related to metabolic traits.
Hfe is also related to iron homeostasis.

Iron metabolism has a tight and complex relationship with diabetes
and obesity [@pmid35466128,@pmid33876811,@pmid32752277]. Adipose
tissue macrophages have systemic effects on metabolism through manipulation
of local iron concentrations [@pmid36329217]. Adding an iron chelator to 
an HFD reduces weight gain in mice compared with HFD alone [@pmid35031684]. 
The mice fed the chelator had more beige fat than their control counterparts.
Isogenic ob/ob mice fed iron in a high-fat diet have a wide range of responses 
[@pmid34646852], suggesting a response more complex than "simple" genetic control.

```{r literature}
library(easyPubMed)

lit.genes <- just.triage[,1]
all.pubmed <- lapply_pb(as.vector(lit.genes), 
    function(x) get_pubmed_ids(paste("Diabetes", "AND ", x)))
num.papers <- sapply(all.pubmed, function(x) x$Count)
with.papers <- which(num.papers != 0)
cbind(lit.genes, num.papers)
cbind(lit.genes[with.papers], num.papers[with.papers])
has.data <- all.pubmed[with.papers]
#all.abstracts <- lapply(has.data, function(x) fetch_pubmed_data(x, format = "abstract"))
#names(all.abstracts) <- lit.genes[with.papers]
#str(all.abstracts)
```

The following genes have been connected to adipose and diabetes in the
literature:

```{r diabetes_genes}
cat(names(all.abstracts), sep = "\n")
```