---
title: "SOM CC-RIX"
author: Anna L Tyler
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    collapsed: no
    toc: yes
    toc_float: yes
---

The purpose of this workflow is examine the use of self-organizing maps 
to analyze transcriptome data for the CC-RIX.


```{r get_args}
args <- commandArgs(trailingOnly = T)
tissue.name <- args[1]
delete_previous <- as.logical(args[2])

if(is.na(tissue.name)){
  tissue.name = "adipose"
  delete_previous <- FALSE
}
  is.interactive <- FALSE
  #is.interactive <- TRUE
```

## SOM Results

```{r source_code}
library("here")
all.fun <- list.files(here("Code"), pattern = ".R", full.names = TRUE)
for(j in 1:length(all.fun)){source(all.fun[j])}
```


```{r load_libraries,  warning = FALSE, message = FALSE, error = FALSE}
needed.packages <- c("oposSOM", "gProfileR", "pheatmap", "qtl2", "grid", "stringr",
  "RColorBrewer")
load_libraries(needed.packages, personal.library = TRUE)
```


Get the expression for the specified tissue.


```{r read_data}
all.var <- ls()
data.loaded <- as.logical(length(which(all.var == "expr")))
data.dir <- here("Data", "CC-RIX")

if(!data.loaded){
    annot <- read.csv(file.path(data.dir, "final_sample_annotations.csv"))
    tissue.expr <- readRDS(file.path(data.dir, "Expression.by.Tissue.RDS"))
    manifest <- as.matrix(read.csv(file.path(data.dir, "Original_Samples_Manifest.csv")))
    covar <- as.matrix(read.csv(file.path(data.dir, "Covar.csv"), row.names = 1))
    bw <- as.matrix(read.csv(file.path(data.dir, "Weight.Summarized.csv"), header = TRUE, row.names = 1))
    chem.mat <- read.csv(file.path(data.dir, "Blood.Chemistries.Summarized.csv"), row.names = 1)
    groups <- read.csv(file.path(data.dir, "Groups.csv"), row.names = 1)
    group.def <- colnames(groups)
}
```

Adjust expression for RNASeq_Batch and Age.
Adjust phenotypes for Age.

We will also compare animals within a
group defined by treatment, diet, and sex using the
contrasts available in oposSOM.

```{r adjust}
tissue.locale <- which(names(tissue.expr) == tissue.name)

orsam.id <- names(rownames(tissue.expr[[tissue.locale]]))
orsam.idx <- match(orsam.id, rownames(covar))

rnaseq.covar <- dummy_covar(covar[orsam.idx,c("Age", "RNAseq_Batch")])
adj.expr <- adjust(tissue.expr[[tissue.locale]], rnaseq.covar)
saveRDS(adj.expr, here("Results", "SOM", paste0("Adjusted.Expression.", tissue.name, ".RDS")))

mouse.info <- unique(t(sapply(rownames(covar), function(x) get_mouse_info(x, manifest, "User.Inputs"))))
rownames(mouse.info) <- mouse.info[,1]

pheno.covar <- dummy_covar(mouse.info[,"Timepoint",drop=FALSE])
num.chem <- apply(as.matrix(chem.mat[,5:ncol(chem.mat)]), 2, as.numeric)
rownames(num.chem) <- rownames(chem.mat)
adj.chem <- adjust(num.chem, pheno.covar)
saveRDS(adj.chem, here("Results", "SOM", "Adjusted.Chem.RDS"))
adj.bw <- adjust(bw, pheno.covar)
saveRDS(adj.bw, here("Results", "SOM", "Adjusted.BW.RDS"))
```

Divide the animals into the groups we will use to study
the specified contrast. If we are looking at the effects
of metformin, we will look at male and female mice with
the high- and low-fat diets separately.

Run the SOM.

```{r som_all_groups}
tissue.info <- unique(t(sapply(rownames(adj.expr), 
  function(x) get_mouse_info(x, manifest, "Barcode.1"))))
all.groups <- apply(tissue.info, 1, function(x) paste(x[3:5], collapse = "_"))

#assign different colors to the groups
u_groups <- unique(all.groups)
group.colors <- brewer.pal(8, "Dark2")
ind.cols <- rep(NA, length(all.groups))
for(i in 1:length(u_groups)){
  ind.cols[which(all.groups == u_groups[i])] <- group.colors[i]
}

som.result.file <- here("Documents", "1.Setup_Data", 
  paste0("CC-RIX_", tissue.name, ".RData"))
stop()
if(delete_previous){
  unlink(som.result.file)
}

#generate groups for comparison
comparison.lists <- list(
    "met_vs_none" = list(
    "met" = c("Female_HFD_Metformin", "Female_LFD_Metformin", "Male_LFD_Metformin", "Male_HFD_Metformin"),
    "none" = c("Female_HFD_None", "Female_LFD_None", "Male_LFD_None", "Male_HFD_None")),
    "sex" = list(
      "female" = c("Female_HFD_None", "Female_HFD_Metformin", "Female_LFD_None", "Female_LFD_Metformin"),
      "male" = c("Male_LFD_None", "Male_HFD_None", "Male_LFD_Metformin", "Male_HFD_Metformin")),
    "diet" = list(
      "LFD" = c("Male_LFD_None", "Female_LFD_None", "Female_LFD_Metformin", "Male_LFD_Metformin"),
      "HFD" = c("Female_HFD_None", "Female_HFD_Metformin", "Male_HFD_None", "Male_HFD_Metformin")))

comp.list <- vector(mode = "list", length = length(comparison.lists))
names(comp.list) <- names(comparison.lists)
for(i in 1:length(comparison.lists)){
  one.comp <- vector(mode = "list", length = 2)
  names(one.comp) <- names(comparison.lists[[i]])
  for(j in 1:2){
    one.comp[[j]] <- rownames(adj.expr)[which(all.groups %in% comparison.lists[[i]][[j]])]
  }
  comp.list[[i]] <- one.comp
}

if(!file.exists(som.result.file)){
  exp.name <- gsub(".RData", "", basename(som.result.file))
  env <- opossom.new(list(dataset.name = exp.name))
  env$indata <- t(adj.expr)
  env$group.labels <- all.groups
  env$group.colors <- ind.colors
  #define specific comparisons
  env$preferences$pairwise.comparison.list <- comp.list
  opossom.run(env)
}
```
